Robust positively invariant sets for state dependent and scaled disturbances

### 2015-03-17 draft.tex
Updated title, abstract and introduction

\usepackage{graphics}
\usepackage{epsfig}
% \usepackage{mathptmx}
% \usepackage{times}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{cite}
\usepackage{lpic}
% \usepackage{blkarray}
\usepackage{mathrsfs}
\usepackage{todonotes}

@@

\begin{document}
  \title{Advances in Robust Positive Invariant Set Computation}

@@

\begin{abstract} 
In this paper we introduce methods of deriving and computing maximal positive robust invariant
sets for linear discrete time systems with additive uncertainty. We consider two different sets of
uncertainties, state dependent uncertainty which allows us to handle multiplicative as well as
linearised systems easily, and scaled sets of uncertainty. We provide the framework for the analytic
treatment of both and illustrate their benefits in two numerical examples.
\end{abstract}

@@

\section{Introduction}
Ever since the introduction of disturbance invariant sets in~\cite{Glover:1971}, the analytic
properties of these sets have been fairly well studied, see~\cite{blanchini:2007}. Here by disturbance
invariant sets we mean a subset of the state constraint set $\mathscr X\subseteq\mathcal X_0\subseteq\mathbb R^n$ 
containing initial conditions for the perturbed linear system
\begin{equation}\label{eq:system:equation}
	x^+ = \Psi x + v,
\end{equation}
where $\Psi$ is Hurwitz, $x,x^+\in\mathcal X_0$ the state and the successor state, and where 
$v\in\mathscr V$ denotes the perturbation, such that the successor state is contained in the 
disturbance invariant set $x^+\in\mathscr X$ for all possible realisations of the 
perturbation $v\in\mathscr V$. That is summarised in the implicit definition
\begin{equation}\label{eq:definition:MRPI:set:state:dependent}
	\mathscr X = \{x:\Psi x + v\in\mathscr X\; \forall v\in\mathscr V\}.
\end{equation}
These sets are known as robust positive invariant (RPI) sets. In many applications we are interested in the largest such
set $\mathcal X^\infty$, i.e. the set containing all other RPI sets $\mathscr X\subseteq\mathcal X^\infty$,
this set is called the maximal robust positive invariant (MRPI) set. These sets are used e.g. as terminal regions in 
robust model predictive control. Even though analytical properties
of these sets were derived early after their introduction, it took more than two decades until algorithms
were introduced to numerically compute MRPI sets~\cite{DeSantis:1994,Kolmanovsky:1995,Blanchini:1994}. 
\note{You commented 'check', I didn't know what that meant.}
Earlier algorithms
did not guarantee finite determinability of $\mathcal X^\infty$, see e.g.~\cite{Blanchini:1990}, or 
did not guarantee to produce the maximal RPI set, e.g.~\cite{Blanchini:1991}. 
%
In the afore mentioned algorithms the perturbation was considered to belong to a fixed compact set, in this paper
we consider sets that vary with parameters. First we consider the set of disturbances to be state dependent, e.g.
\begin{equation}\label{eq:PWA:distrubance:set}
	\mathcal V(x) = \left\{v: Gv\leq H(x)\right\},
\end{equation}
where $H(x)$ is a piecewise affine function. This means that in the first part of the paper
we discuss the computation of $\mathcal X^\infty$ for the case where $\mathscr V=\mathcal V(x)$.
The case where the disturbance depends on the state has been studied analytically before but not 
extensively, see~\cite{Kuntsevich:1995}. These kind of disturbance constraints can be used to
account for linearisation errors, as will be shown in section~\ref{sec:first:example}, but can also
be used to treat multiplicative uncertainty.

The second case we consider is the case of a scaled set of disturbance, i.e.
\begin{equation}
  \mathcal V(\theta) = \{v:Gv\leq(1+\theta){\bf{1}}\},
\end{equation}
with $\theta>-1$. To distinguish the resulting sets from regular MRPI sets we use $\mathcal Z^\infty$
to emphasise that $\mathcal Z^\infty$ itself is not an MRPI set, however by fixing $\theta=\hat\theta$ the
set $\mathcal Z^\infty\vert_{\hat\theta}$ is an MRPI set. To our best knowledge, this setup has not been 
presented in the literature. This type of disturbance constraints can be used to study the sensitivity of
the MRPI set to changes in the disturbance \emph{strength}.

The remainder of this paper is structured: In section~\ref{sec:parametrically:convex:set:operations} we
introduce the notion of \emph{parametric convexity} and of its general implications, this is sufficient 
for the convexity of the MRPI set. The algorithm to compute the MRPI set for state dependent disturbance
constraints is presented in section~\ref{sec:state:dep:MRPI}, we also prove its finite determinability.
In section~\ref{sec:first:example} we demonstrate the algorithm using a numerical exemplary system, a
levitating ball setup. The case of scaled disturbance constraints is discussed in section~\ref{sec:MRPI:parametrised}
where the algorithm for the parametrised MRPI set as well as the proof for its finite determinability
are also presented. Section~\ref{sec:second:example} illustrates the use of parametrised MRPI sets using
again the levitating ball example and illustrating some of possible robustness analyses of the parametrised
MRPI sets with a numerical example. The paper is concluded in section~\ref{sec:conclusions}.

In the remainder of this paper we use the notion of a polyhedra for sets that can be represented as
the intersection of a finite number of halfspaces, polytopic sets for bounded polyhedral.
By ${\bf{1}}$ we denote the column vector of ones in appropriate dimension, $\wedge$ denotes the logical and. 

@@

\begin{lpic}{levitatingBall}
\lbl[tr]{25,3; $m g$}
\lbl[br]{25,25; $c\frac{i^2}{x^2}$}
\lbl[bl]{49,17; $y$}
\lbl[bl]{56,55; $i$}

@@

\note{You wrote to highlight $\mathcal V(0)$ being bounded was an assumption. It is necessary so I 
don't know what to do with this.}

@@

The proof is, completely analogue to the proof of Lemma~\ref{thm:finite:MRPI:set:state:dependable}.

@@

\includegraphics[width=\columnwidth]{twoDimensional}

@@

\includegraphics[width=\columnwidth]{threeDimensional}

>>>>

\usepackage{graphics}
\usepackage{epsfig}
% \usepackage{mathptmx}
% \usepackage{times}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{cite}
\usepackage{lpic}
\usepackage{overpic}
% \usepackage{blkarray}
\usepackage{mathrsfs}
\usepackage{todonotes}

@@

\begin{document}
%  \title{Advances in Robust Positively Invariant Set Computation}
  \title{Robust positively invariant sets for state dependent and scaled disturbances}

@@

\begin{abstract} 
  This paper introduces methods of deriving and computing maximal robust positively invariant sets for linear discrete time systems with additive model uncertainty. Two types of uncertainty are considered: state dependent uncertainty, which handles multiplicative parametric model uncertainty as well as linearisation errors for nonlinear systems, and scaled sets of uncertainty. We provide a framework for analysing both types of uncertainty with illustrative examples.
\end{abstract}

@@

\section{Introduction}
The analytic properties of disturbance invariant sets have been widely studied (see e.g.~\cite{blanchini:2007}) since their introduction in~\cite{Glover:1971}. A disturbance invariant set $\mathscr X$ is a subset of the state constraint set $\mathcal X_0\subseteq\mathbb R^n$ that contains states $x$ of the perturbed linear system
\begin{equation}\label{eq:system:equation}
	x^+ = \Psi x + v
\end{equation}
(where $\Psi$ is Hurwitz, $x,x^+\in\mathcal X_0$, and 
$v\in\mathscr V$) such that the successor state $x^+$ is contained in the 
disturbance invariant set for all possible realisations of the 
unknown disturbance $v$. This is summarised in the implicit definition
\begin{equation}\label{eq:definition:MRPI:set:state:dependent}
	\mathscr X = \{x:\Psi x + v\in\mathscr X\; \forall v\in\mathscr V\}.
\end{equation}
Thus an disturbance invariant set is a type of robust positive invariant (RPI) set. In many applications we are interested in the largest such
set, denoted $\mathcal X^\infty$, which is given by the union of all RPI sets $\mathscr X\subseteq\mathcal X^\infty$. This set is known as the maximal robust positive invariant (MRPI) set and has been used extensively, for example to define terminal regions in 
robust model predictive control. Although analytical properties
of these sets were derived soon after their introduction, algorithms for numerically computing MRPI sets were introduced much later~\cite{DeSantis:1994,Kolmanovsky:1995,Blanchini:1994}. 
%\note{You commented 'check', I didn't know what that meant.}
Earlier algorithms
did not guarantee finite determinability of $\mathcal X^\infty$, see e.g.~\cite{Blanchini:1990}, or 
did not guarantee to produce the maximal RPI set, e.g.~\cite{Blanchini:1991}. 
%
However for the case of disturbances that belong to a fixed set $\mathscr V$, methods of computing $\mathcal X^\infty$ are well established~\cite{blanchini:2007}.

This paper considers
sets of disturbances that depend on parameters. First we consider disturbance sets that are state dependent, of the form
\begin{equation}\label{eq:PWA:distrubance:set}
	\mathcal V(x) = \left\{v: Gv\leq H(x)\right\},
\end{equation}
where $H(x)$ is a piecewise affine function. For the case that $\mathscr V=\mathcal V(x)$ we
provide conditions for convexity and finite determination of $\mathcal X^\infty$ and discuss computation of $\mathcal X^\infty$.
The case of disturbances depending on states (or control inputs) is considered in~\cite{Kuntsevich:1995,rakovic06}, but the convexity analysis of this paper is novel. State dependent disturbance constraints can 
account for linearisation errors, as shown in section~\ref{sec:first:example}, as well as more general multiplicative parametric model uncertainty.

The second case considered is that of scaled disturbance sets of the form
\begin{equation}
  \mathcal V(\theta) = \{v:Gv\leq(1+\theta){\bf{1}}\},
\end{equation}
for scalar $\theta>-1$. For $\mathscr V = \mathcal V(\theta)$ we compute sets $\mathcal Z^\infty$ in $(x\times\theta)$-space with the property that 
% To distinguish the resulting sets from regular MRPI sets we use $\mathcal Z^\infty$
% to emphasise that $\mathcal Z^\infty$ 
the intersection, $\mathcal Z^\infty\vert_{\hat\theta}$, with the subspace on which $\theta=\hat\theta$ for given $\hat\theta$ is an MRPI set. To our best knowledge, this setup has not been 
considered in the literature. This type of disturbance constraint can be used to study the sensitivity of
the MRPI set to changes in the disturbance \emph{strength}.

The paper is structured as follows: In section~\ref{sec:parametrically:convex:set:operations} we
introduce the notion of \emph{parametric convexity} and discuss its implications, in particular as a condition 
for the convexity of the MRPI set. The algorithm to compute the MRPI set for state dependent disturbance
constraints is presented in section~\ref{sec:state:dep:MRPI}, and we also prove its finite determinability.
In section~\ref{sec:first:example} the algorithm is illustrated using an exemplar system, a magnetically levitated ball. The case of scaled disturbance constraints is discussed in section~\ref{sec:MRPI:parametrised}
where the algorithm for the parametrised MRPI set as well as the proof for its finite determinability
are also presented. Section~\ref{sec:second:example} illustrates the use of parametrised MRPI sets using
again the levitating ball example and illustrating a robustness analyses of the parametrised
MRPI sets with a numerical example. The paper is concluded in section~\ref{sec:conclusions}.

Throughout this paper we refer to sets that can be represented as intersections of finite numbers of half spaces as polyhedra, and to bounded polyhedra as polytopic sets.
%By ${\bf{1}}$ we denote 
Also ${\bf{1}}$ denotes the column vector of ones in appropriate dimensions, and $\wedge$ is the logical AND operator. 

@@

\begin{overpic}[scale=0.75]{levitatingBall}
\put(30,5){$m g$}
\put(30,41){$c\frac{i^2}{x^2}$}
\put(78,28){$y$}
\put(90,89){$i$}
\end{overpic}
% \begin{lpic}[scale=0.5]{levitatingBall}
% \lbl[tr]{25,3; $m g$}
% \lbl[br]{25,25; $c\frac{i^2}{x^2}$}
% \lbl[bl]{49,17; $y$}
% \lbl[bl]{56,55; $i$}
% \end{lpic}
\vspace{-2mm}

@@

%\note{You wrote to highlight $\mathcal V(0)$ being bounded was an assumption. It is necessary so I don't know what to do with this.}

@@

The proof is similar to that of Lemma~\ref{thm:finite:MRPI:set:state:dependable}.

@@

\centerline{\includegraphics[scale=0.6]{twoDimensional}}

@@

\centerline{\includegraphics[scale=0.6]{threeDimensional}}
